Staged Abstract Interpreters

### 2019-09-07 paper/10_intro.tex
intro

In this paper, we show that the first Futamura projection can be naturally
extended to abstract interpreters. We present an abstraction-without-regret
framework that eliminates performance penalties for monadic abstract
interpreters, while keeping the implementation as close to their conceptual
model as possible.
In short, our approach borrows ideas and techniques from multi-stage
programming and embedded DSLs, and apply them to abstract interpreters:  1) To
remove the overhead from interpretation and effect layers, we specialize the
abstract interpreter via staging and then generate efficient low-level code
that does the actual analysis. 2) Inspired by the tagless-final interpreters
\cite{DBLP:journals/jfp/CaretteKS09}, we construct a generic interpreter that
abstracts over binding-times and different semantics, which allows the staged
abstract interpreter to be derived from its unstaged counterpart.  As a result,
the derived staged abstract interpreters has no intrusive changes to the
underlying abstract semantics, thereby preserving soundness.  In this
sense, our approach allows for no regret for both performance and engineering
effort. We elaborate these two main ideas in detail, as follows.

>>>>

In this paper, we show that the first Futamura projection can be naturally
extended to abstract interpreters. We present an abstraction-without-regret
framework that eliminates performance penalties for monadic abstract
interpreters, while keeping the implementation as close to their conceptual
model as possible.
In short, we borrow ideas and techniques from multi-stage programming and
embedded DSLs, and apply them to abstract interpreters:  1) To
remove the overhead from interpretation and effect layers, we specialize the
abstract interpreter via staging and then generate efficient low-level code
that performs the actual analysis. 2) Inspired by the tagless-final interpreters
\cite{DBLP:journals/jfp/CaretteKS09}, we construct a generic interpreter that
abstracts over binding-times and different semantics, which allows the staged
abstract interpreter to be derived from its unstaged counterpart.  As a result,
the derived staged abstract interpreters has no intrusive changes to the
underlying abstract semantics, thereby preserving soundness.  In this
sense, our approach allows for no regret for both performance and engineering
effort. We elaborate these two main ideas in detail, as follows.
